---
title: "Assignment 3"
author: "Pierre-Olivier Couture"
format: html
editor: visual
---

## Assignment 3

This is all the code I used for Assignment 3

```{r}
#| label: Packages
# This section of code installs and initializes all the packages I will use throughout the assignment. The flexplot packages was "#" out because it kept asking me for updates and will not let me render into a html file. The code still works fine with this adjustment since the package is already installed on R. 
library(lme4)
options(repos = c(CRAN = "https://cran.rstudio.com/"))

#install.packages("remotes")

#remotes::install_github("dustinfife/flexplot")

install.packages("apaTables")
install.packages("broom.mixed")

library(tidyverse)
library(haven)      
library(lme4)       
library(lmerTest)   
library(dplyr)
library(performance) 
library(sjPlot)     
library(psych)    

library(haven)
library(tidyverse)
library(dplyr)
library(flexplot)
library(apaTables)

```

```{r}
#| label: Load in the data
# This section of code loads in the data and cleans the names to remove all capital letters to make it easier when I need to call on the variables throughout my code. 

A4 <- read_sav("~/Desktop/P6003.A4.sav")
View(A4)

```

```{r}
#| label: Correlation table
# I used this section of code to create a data frame that includes all the variables and makes it easier for me to report the descriptive statistics. Could include them in a table if I have time to look at the correct code to do this since the apaTables package does not have a code to report descriptive statistics and I would need to find more background information.  

describe_vars <- A4 %>% select(swl, tipm.E, tipm.N)
psych::describe(describe_vars)

# This part of the code will provide me with the correlation between all the variables using only the rows with no missing data for the variables I am interested in. This will only not include variables if any of the variables have missing data for the variables of interest.

cor(describe_vars, use = "pairwise")
apa.cor.table(describe_vars, filename = "Correlation Matrix of Satisfaction with Life, Extraversion, and Neuroticism.doc", table.number = 1)


```


```{r}
#| label: H1 Code
# This is the code I will use to test all assumptions and calculate all the data needed for H1. 



```
The code above does an assumption check for H1 and creates a model for H1. Below is my interpretation of everything to help me with the write-up later.  

-- The Histogram of Residuals shows that the residuals are approximately normally distributed but with a slight left skew.If this is the only violation of the assumptions, the model is still considered reliable. However, if other assumptions are also violated, the model may not be valid. NORMALITY of RESIDUALS assumption is met. 

-- The RDP is flat meaning this assumption is met. LINEARITY assumption is met.

-- The S-L Plot is relatively flat and straight meaning this assumption is met. HOMOSKEDACITY assumption is met.

-- The ACF plot shows that several of the lines extend past the 95% CI range meaning that the residuals at one time point seem to be related to the residuals from previous time points. The assumption is not met. INDEPENDENCE of OBSERVATIONS assumption NOT met.

-- Multicolinearity was not assessed since I was only using 1 predictor and this cannot be calculated. So multicolinearity is not an issue in H1. 

All I need to do is mention that the independence of observations assumption is violated and that I would conduct a mixed-effects model to model the temporal dependencies between observations. *** ASKED ASHLEY IF THIS NEEDED TO BE INCLUDED OR IF I CAN JUST MENTION IT ***

Using the estimate and summary function here are the findings to make it easier to write about in the Word doc.

Estimate for tipm.E = 0.40 meaning the association is positive and this is considered a moderate association.
-- The 95% CI are 0.37 and 0.43. Since these do not cross 0 we can say this association is statistically significant. The CI also have a small margin of error so we can say the estimate is relatively precise. 
-- The standardized estimate is 0.38 with the CIs of 0.35 and 0.41 further supporting that there is a moderate, positive association between extraversion and satisfaction with life
-- Multiple R^2 is 0.1441 and the adjusted R^2 is 0.1439.
---- Multiple R^2 suggests that 14.41% of the variance in satisfaction with life is explained by extraversion. Given this number is quite small, we can assume that other predictors influence this association. 
---- Semi-Partial R^2 indicates the unique proportion of variance in satisfaction with life that is explained by extraversion alone, excluding other potential predictors. 0.144 suggests a moderate, positive relationship between extraversion and satisfaction with life, meaning extraversion has a meaningful effect on satisfaction with life. 

```{r}
#| label: H2 Code
# This is the code I will use to test all assumptions and calculate all the data needed for H2.

# Step 1: Creating the regression model for H2
H2 <- lm(swl ~ tipm.N, data = A4)

# Step 2: Visualize the model
visualize(H2, plot = "residuals")

# Step 3: Model Output for H2
estimates(H2)
summary(H2)

# Step 4: Extract residuals for further assumption checks
residuals_H2 <- residuals(H2)

# Step 5: Check for independence of observations (ACF plot for residuals)
# Using this to examine the independence of observations since the data was collected over several time points.
acf(residuals_H2)

# Step 6: Check normality of residuals (Q-Q plot)
# Use Q-Q plot to visually check if residuals follow a normal distribution
qqnorm(residuals_H2)
qqline(residuals_H2, col = "red")

# Step 7: Mixed-effects model to account for non-independence of observations
# Random intercept for id (subject), fixed effect for tipm.N (neuroticism), and factor(day)
H2_mixed <- lmer(swl ~ tipm.N + factor(day) + (1 | id), data = A4)

# Step 8: Display the summary of the mixed-effects model
summary(H2_mixed)

# Step 9: Check residuals of the mixed-effects model
# Plot residuals of the mixed-effects model to check for patterns or outliers
residuals_H2_mixed <- residuals(H2_mixed)
plot(residuals_H2_mixed)

# Step 10: Check random effects
# Random intercept variance for id
ranef(H2_mixed)

# Step 11: Check confidence intervals for mixed model coefficients
# To get confidence intervals for the fixed effects
confint(H2_mixed)

# Step 12: Intraclass Correlation (ICC) to see the proportion of variance explained by random effects
icc(H2_mixed)

# Step 13: R-squared for the mixed-effects model (marginal and conditional R²)
r2(H2_mixed)

# Display the mixed-effects model output in a table format
tab_model(H2_mixed, show.ci = TRUE, show.re.var = TRUE, show.icc = TRUE, show.r2 = TRUE)

```
The code above does an assumption check for H2 and creates a model for H1. Below is my interpretation of everything to help me with the write-up later.  

-- The Histogram of Residuals shows that the residuals are approximately normally distributed but with a slight left skew.If this is the only violation of the assumptions, the model is still considered reliable. However, if other assumptions are also violated, the model may not be valid. NORMALITY of RESIDUALS assumption is met. 

-- The RDP is is slightly curved at the both ends with a tip about 3/4 of the way into the line. From the previous assignment Ashley said you're good to just mention it appears to be violated and that a quadratic or other model may be a better fit and move along as if the assumption was met! However, make sure to include in the results that results need to be interpreted with caution given the issue with linearity. LINEARITY assumption NOT met. 

-- The S-L Plot is relatively flat and straight meaning this assumption is met. HOMOSKEDACITY assumption is met.

-- The ACF plot shows that several of the lines extend past the 95% CI range meaning that the residuals at one time point seem to be related to the residuals from previous time points. The assumption is not met. INDEPENDENCE of OBSERVATIONS assumption NOT met.

-- Multicolinearity was not assessed since I was only using 1 predictor and this cannot be calculated. So multicolinearity is not an issue in H1. 

All I need to do is mention that the independence of observations assumption is violated and that I would conduct a mixed-effects model to model the temporal dependencies between observations. *** ASKED ASHLEY IF THIS NEEDED TO BE INCLUDED OR IF I CAN JUST MENTION IT ***

Using the estimate and summary function here are the findings to make it easier to write about in the Word doc.

Estimate for tipm.N = -0.47, meaning the association is negative and this is considered a moderate association. 
-- The 95% CI are -0.50 and -0.44. Since these do not cross 0, we can say this association is statistically significant. The CI also has a small margin of error, so we can say the estimate is relatively precise. 
-- The standardized estimate is -0.45 with the CIs of -0.48 and -0.42, further supporting that there is a moderate, negative association between neuroticism and satisfaction with life.
-- Multiple R² is 0.2035, and the adjusted R² is 0.2033. 
---- Multiple R² suggests that 20.35% of the variance in satisfaction with life is explained by neuroticism. Given this number is quite small, we can assume that other predictors influence this association. 
---- Semi-Partial R² indicates the unique proportion of variance in satisfaction with life that is explained by neuroticism alone, excluding other potential predictors. 0.144 suggests a moderate, negative relationship between neuroticism and satisfaction with life, meaning neuroticism has a meaningful effect on satisfaction with life.

```{r}
# Full model with both predictors
H3_model <- lmer(
  swl ~ extraversion_mean + extraversion_wp + neuroticism_mean + neuroticism_wp + (1 | id),
  data = A4,
  REML = FALSE
)

summary(H3_model)
tab_model(H3_model, show.ci = TRUE, show.re.var = TRUE)
r2(H3_model)

# Test random slopes to see if effects vary by person
H3_model_random_slopes <- lmer(
  swl ~ extraversion_mean + extraversion_wp + neuroticism_mean + neuroticism_wp +
    (1 + extraversion_wp + neuroticism_wp | id),
  data = A4,
  REML = FALSE
)

anova(H3_model, H3_model_random_slopes)  # Compare fixed vs random slopes

```



```{r}
#| label: H3 Code
# This is the code I will use to test all assumptions and calculate all the data needed for H3.

# Step 1: Creating the multilevel regression model with both Extraversion (tipm.E) and Neuroticism (tipm.N)
H3 <- lmer(swl ~ tipm.E + tipm.N + (1 | id), data = A4)

# Step 2: Visualize the residuals for the model
visualize(H3, plot = "residuals")

# Step 2.5: Checking for multicolinearity
lm_model <- lm(swl ~ tipm.E + tipm.N, data = A4)

# Calculate VIFs
vif(lm_model)

# Step 3: Model Output for H3
summary(H3)

# Step 4: Extract residuals for further assumption checks
residuals_H3 <- residuals(H3)

# Step 5: Check for independence of observations (ACF plot for residuals)
# Using this to examine the independence of observations since the data was collected over several time points.
acf(residuals_H3)

# Step 6: Visualize the relationship between swl and both tipm.E and tipm.N using flexplot
# Since I have checked the assumptions above, this code will provide me with a linear regression model of the relationship between swl, tipm.E, and tipm.N and allows me to visualize the relationship.
flexplot(swl ~ tipm.E + tipm.N, data = A4)

# Step 7: Provide CI
confint(H3, method = "profile")

apa.reg.table(H3, conf.level = 0.95)

```
The code above does an assumption check for H3 and creates a model for H1. Below is my interpretation of everything to help me with the write-up later.  

-- The Histogram of Residuals shows that the residuals are normally distributed around 0. NORMALITY of RESIDUALS assumption is met. 

-- The RDP is is slightly curved at the both ends. From the previous assignment Ashley said you're good to just mention it appears to be violated and that a quadratic or other model may be a better fit and move along as if the assumption was met! However, make sure to include in the results that results need to be interpreted with caution given the issue with linearity. LINEARITY assumption NOT met. 

-- The S-L Plot is relatively flat and straight meaning this assumption is met. HOMOSKEDACITY assumption is met.

-- The ACF plot shows that the majority of the lines fall within the CI lines meaning the answers from previous time points do not seem to significantly impact the scores at future time points. ASSUMPTION of INDEPENDENCE is met. 

-- Multicolinearity was assessed using vif found that both tipm.E and tipm.N have low multicolinearity (1.124797 and 1.124797, respectively). Any value below 5 is considered to mean the assumption is met and given that our values are ~1 we can say this assumption is met. MULTICOLINEARITY assumption is met. 

NEUROTICISM
- The estimate for tipm.N is -0.23, indicating a moderate, negative association with satisfaction with life.

- The 95% confidence interval for tipm.N is [-0.25, -0.21], which means we are 95% confident that the true relationship between neuroticism and life satisfaction lies within this range. Since the interval does not include 0, we can confidently conclude that neuroticism is negatively associated with life satisfaction.

-- Individuals higher in neuroticism tend to report lower satisfaction with life.

EXTRAVERSION
- The estimate for tipm.E is 0.16, indicating a moderate, positive association with satisfaction with life.

- The 95% confidence interval for tipm.E is [0.13, 0.18], meaning we are 95% confident that the true relationship between extraversion and life satisfaction lies within this range. Since the interval does not include 0, we can confidently conclude that extraversion is positively associated with life satisfaction.

-- Individuals higher in extraversion tend to report higher satisfaction with life.

RANDOM EFFECTS
Random intercept variance (for id): The 95% confidence interval for the random intercept variance is [1.10, 1.32], suggesting that the variability in life satisfaction between individuals is substantial.

Residual variance: The 95% confidence interval for the residual variance is [0.73, 0.76], indicating that within-person variability in life satisfaction is moderate.

These random effects indicate that there is a significant amount of individual variability in life satisfaction (both between individuals and within individuals), underscoring the need for a multilevel model.









The `echo: false` option disables the printing of code (only output is displayed).
